---
title: "R Programming: <br>Replacing For Loops with lapply"
subtitle: "when Creating Several Plots"
author: "StatistikinDD"
date: "Presentation Created: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["xaringan-themer.css", "libs/_css/my_css.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%"
      ratio: 16:9
---

```{r setup, include = FALSE}

options(htmltools.dir.version = FALSE)
# knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, comment = "")
knitr::opts_chunk$set(echo = TRUE, comment = "")
xaringanthemer::style_mono_accent(title_slide_background_color = "#EDF4FB",
                                  title_slide_text_color = "#000000",
                                  base_font_size = "18px")
Sys.setenv(LANG = "en")

# Trick to suppress lapply's list output
# Source: https://stackoverflow.com/questions/54300147/suppress-list-notation-in-rmarkdown-from-lapply

def <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
  x <- def(x, options)
  ifelse(!is.null(options$suppress), gsub(pattern = "```.*```", "", x), x)
})

```

# The Data

## mpg {ggplot2}: Fuel economy data from 1999 to 2008 for 38 popular models of cars

```{r mpg-data, message = FALSE}

library(tidyverse)
glimpse(mpg)

```

---

# The Data

## Which Manufacturers?

.pull-left[
```{r mpg-manuf, eval = FALSE}

theme_set(ggthemes::theme_gdocs(base_size = 14))

mpg <- mpg %>%
  mutate(manuf_c = fct_lump_min(manufacturer,#<<
                                min = 20))

ggplot(mpg, aes(y = fct_rev(
                fct_infreq(manufacturer)))) +
  geom_bar(aes(fill = manuf_c)) +
  scale_fill_manual(values = c( #<<
        scales::hue_pal()(4), "grey50"),
        name = "") + #<<
  labs(title = "Manufacturers in the mpg Dataset",
       x = "Count", y = "Manufacturer")

```

]

.pull-right[
```{r, ref.label = "mpg-manuf", echo = FALSE}
```

]

---

# Fuel Efficiency (City Miles per Gallon) <br>vs. Engine Displacement

.pull-left[

```{r plot, eval = FALSE}

ggplot(mpg, aes(x = displ, y = cty,
                color = manuf_c)) +
  geom_point(size = 3) +
  scale_color_manual(values = c(
        scales::hue_pal()(4), "grey50"),
        name = "Manufacturer") +
  labs(title = "Fuel Efficiency vs. Engine Displacement",
       x = "Displacement", y = "City Miles per Gallon")

```

* Not so easy to distinguish manufacturers
* Could use facets, but here we want separate plots

]

.pull-right[
```{r, ref.label = "plot", echo = FALSE}
```

]

---

# One Plot per Manufacturer: A For Loop?

.pull-left[
```{r loop, eval = FALSE}

for(i in levels(mpg$manuf_c)) { #<<
  p <- mpg %>%  #<<
    filter(manuf_c == i) %>% 
    ggplot(aes(x = displ, y = cty)) +
      geom_point(size = 3) +
      geom_smooth(method = "lm", color = "blue") +
      labs(title = "Fuel Eff vs. Displmt",
           subtitle = paste("Manufacturer:", i),
           x = "Displacement", y = "City Miles per Gallon") +
    ggthemes::theme_gdocs(base_size = 10)
    print(p)  #<<
    # ggsave(p, filename = paste0("mpg_", i, ".png"))
} #<<

```
]

.pull-right[
```{r, ref.label = "loop", fig.show = "hold", echo = FALSE, fig.width = 3, fig.height = 3, message = FALSE}
```

]

---

# A Functional Programming Approach

.pull-left[

### Defining Our Plotting Function

```{r function-def}

my_plot <- function(manuf) { #<<
  mpg %>%
    filter(manuf_c == manuf) %>% #<<
    ggplot(aes(x = displ, y = cty)) +
      geom_point(size = 3) +
      geom_smooth(method = "lm", color = "blue") +
      labs(title = "Fuel Eff vs. Displmt",
           subtitle = paste("Manufacturer:", manuf)) + #<<
    ggthemes::theme_gdocs(base_size = 10)
}

```
]

--

.pull-right[
### Applying the Function

```{r function-app, fig.show = "hold", fig.width = 3, fig.height = 3, message = FALSE}

my_plot("dodge")
my_plot("Other")

```

]

---

# lapply: A One-Liner to Avoid<br>Copy & Paste & loops

.pull-left[
```{r lapply, eval = FALSE}

manufs <- levels(mpg$manuf_c)

lapply(manufs, my_plot) #<<

```

* No print statement required

* If you want to save plots, you can assign to an object (p) and use  
  `ggsave(p, file = ...)`  
  inside the user-defined function

]

.pull-right[
```{r, ref.label = "lapply", echo = FALSE, fig.show = "hold", fig.width = 3, fig.height = 3, message = FALSE, suppress = TRUE}
```

]

---

# Last Words

### lapply() extends well to advanced use cases
* **See my videos on running R code in parallel: *parallel::clusterApply() / clusterApplyLB()* **
* **Processing large datasets in chunks: *iotools::chunk.apply()* **

--

### Loops can be run in parallel using *foreach* and *%dopar%*

--

### Note that lapply() returns a list
* **Output like [[1]] ... [[2]] ... [[3]] ... etc. in console / Markown document**
* **Suppressed here for sake of presentation**
* **Use *purrr::walk()* instead of lapply to return plots silently**

--

### Possible to avoid explicit function assignment
* **Create anonymous (unnamed, unassigned) function inside lapply**

---

class: center, middle

# Thanks!

### Youtube: StatistikinDD

### Twitter: @StatistikinDD

### github: fjodor

Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](https://yihui.org/knitr), and [R Markdown](https://rmarkdown.rstudio.com).

Thanks to **Yihui Xie** for *{knitr}* and *{xaringan}* and **Garrick Aden-Buie** for *{xaringanthemer}*.